

\chapter{14. The Paired $t-$Test}




\section{Two Sample Inference Procedures}
\begin{itemize}
\item Previously we looked at inference procedures (Confidence Intervals and Hypothesis Testing) for single samples.
\item Yesterday we looked at \textit{\textbf{paired}} samples, with two sets of paired measurements. With paired measurements, we are specifically interested in the \textbf{\textit{case-wise}} differences.
\item Although there are two sets of data, we consider the single data set of case-wise differences.
\item Now we look at the case of two independent sample procedures.
\item Independent samples are distinct from paired samples, in that data in one set are not paired with data in another set.
\end{itemize}




\subsection{Mean Difference Between Matched Data Pairs}


The approach described in this lesson is valid whenever the following conditions are met:

\begin{itemize}
\item The data set is a simple random sample of observations from the population of interest.
\item Each element of the population includes measurements on two paired variables (e.g., x and y) such that the paired difference between x and y is: d = x - y.
\item The sampling distribution of the mean difference between data pairs (d) is approximately normally distributed.
\end{itemize}



The observed data are from the same subject or from a matched subject and are drawn from a population with a normal distribution
does not assume that the variance of both populations are equal.

%---------------------------------------------------------%







\section{The Paired $t-$Test}



\begin{framed}
The standard error of the mean difference between paired
observations is obtained for the standard error of the mean.
\section{Hypotheses}
\begin{eqnarray*}
H_{0}: \mu_{d} = 0\\
H_{1}: \mu_{d} \neq 0\\
\end{eqnarray*}
\end{framed}



\section{Paired T test}
\begin{itemize} \item A paired sample t-test is used to determine whether there is a significant difference between the average values of the same measurement made under two different conditions. \item Both measurements are made on each unit in a sample, and the test is based on the paired differences between these two values. \item The usual null hypothesis is that the difference in the mean values is zero. For example, the yield of two strains of barley is measured in successive years in twenty different plots of agricultural land (the units) to investigate whether one crop gives a significantly greater yield than the other, on average.
\end{itemize}

\textbf{Confidence Interval}:
\begin{itemize}
\item Recall our sample mean $\bar{x}$, the standard error $S.E(\bar{x})$ and the quantile from the $t-$ distribution.
\item We can use these values to compute a 95\% confidence interval.
\item The 95\% confidence interval can be computed as $0.17 \pm (2.262 \times 0.17 = (-0.21,0.55)$
\item Notice that 0 is within that range of values. This supports our conclusion to reject the Null Hypothesis.
\end{itemize}


%----------------------------------------------------------------------------------------------------%

\textbf{Paired t test}
The paired t test was devloped by Guinness Brewery employee, Gossett, in1908.
\begin{itemize}
\item First we have to compute the case-wise differences.
\item Then we compute the variance of those differences.
\end{itemize}






%---------------------------------------------------------%

\[ ( \bar{X} - \bar{Y} ) \pm \left[ \mbox{Quantile } \times S.E(\bar{X}-\bar{Y}) \right] \]
\begin{itemize}
\item If the combined sample size of X and Y is greater than 30, even if the individual sample sizes are less than 30, then we consider it to be a large sample.
\item The quantile is calculated according to the procedure we met in the previous class.
\end{itemize}

%---------------------------------------------------------%
\begin{itemize}
\item Assume that the mean ($\mu$) and the variance ($\sigma$) of the distribution
of people taking the drug are 50 and 25 respectively and that the mean ($\mu$)
and the variance ($\sigma$) of the distribution of people not taking the drug are
40 and 24 respectively.
\end{itemize}









\section{Paired t test}

{
\subsection{Inference : Paired values}

\begin{itemize}

\item Know how to compute case-wise differences.
\item Know how to compute the mean of the case-wise differences (see formulae). 
\item Know how to compute the standard deviation of the casewise differences (see formulae).
\end{itemize} 
}



\begin{itemize}
\item Consider two populations X and Y that are indepedently distributed from
each other.
\item That is to say, the true value of correlation is zero.
\[\rho_{XY} = 0 \]
\item In the context of a linear regression model, in the form $Y=\beta_0  +  \beta_1X$, a true correlation value of zero is equivalent to a true slope value of Zero.
\[``\rho_{XY} = 0" \leftarrow\rightarrow ``\beta_1=0"\]
\end{itemize}

%----------------------------------------%



\subsection{Paired t-test}

%-----------------------------------------------------------%








\[ ( \bar{X} - \bar{Y} ) \pm \left[ \mbox{Quantile } \times S.E(\bar{X}-\bar{Y}) \right] \]
\begin{itemize}
\item If the combined sample size of X and Y is greater than 30, even if the individual sample sizes are less than 30, then we consider it to be a large sample.
\item The quantile is calculated according to the procedure we met in the previous class.


\end{itemize}





%---------------------------------------------------------%



\section{Confidence interval for the true mean difference}
The in above example the estimated average improvement is just over 2 points. Note that
although this is statistically significant, it is actually quite a small increase. It would be
useful to calculate a confidence interval for the mean difference to tell us within what limits
the true difference is likely to lie. 



Using our example:
We have a mean difference of 2.05. The 2.5\% point of the t-distribution with 19 degrees
of freedom is 2.093. The 95\% confidence interval for the true mean difference is therefore:
$2.05 \pm (2.093 \times 0.634) = 2.05 \pm 1.33 = (0.72, 3.38)$.

This confirms that, although the difference in scores is statistically significant, it is actually
relatively small. We can be 95\% sure that the true mean increase lies somewhere between
just under one point and just over 3 points.









%---------------------------------------------------------%

\section{Difference in Two means}
For this calculation, we will assume that the variances in each of the two populations are equal. This assumption is called the assumption of homogeneity of variance.

The first step is to compute the estimate of the standard error of the difference between means ().

\[ S.E.(\bar{X}-\bar{Y}) = \sqrt{\frac{s^2_x}{n_x} + \frac{s^2_y}{n_y}} \]

\begin{itemize}
\item $s^2_x$ and $s^2_x$ is the variance of both samples.
\item $n_x$ and $n_y$ is the sample size of both samples.
\end{itemize}
The degrees of freedom is $n_x + n_y -2$.




%%--------------------------------------------------------------------------------------%
%
%\textbf{Computing the Test Statistic}
%
%The general structure of a test statistic is as follows:
%\[ TS = {\mbox{observed value} - \mbox{null value} \over \mbox{Std. Error}}   \]
%where ``null value" is shorthand for the expected value under the null hypothesis.
%
%Refer to the formulae for the appropriate standard error.
%




%----------------------------------------------------------------------------------%

\section{Worked Examples} 

Two procedures for sintering copper are compared by testing each procedure on six different types of powder. The measurement of interest is the porosity of each test specimen.
The results of the test are as follows:  

\[
\begin{array}{|c|c|c|}
Powder&Procedure 1& Procedure 2\\\hline
1&21&23\\\hline
2&27&26\\\hline
3&18&21\\\hline
4&22&24\\\hline
5&26&25\\\hline
6&19&16\\\hline
\end{array} 
\]




ls there a difference between the true average porosity measurements for the two procedures, at a significance level of 5\%.


%-------------------------------------------------------------------------------------------%



\section{Worked Examples : Paired T test}
%Question 4 part a - 
\begin{itemize}
\item The weight of 6 individuals (in kgs) was observed before and after a diet regime (diet given below)
\item Compute the mean difference and standard deviation of the differences.

\item Remark : Sample size n= 6
\end{itemize}

\[
\begin{array}{cccc}
Person&Weight Before&Weight After&Difference\\ \hline
A&89&87&2\\ \hline
B&79&76&3\\ \hline
C&106&105&1\\ \hline
D&92&92&0\\ \hline
E&88&84&4\\ \hline
F&98&96&2\\ \hline
\end{array} 
\]

Before we start, we need to compute the average difference and the standard deviations of the differences. 



%Sample Mean:           x=xin      


What is the mean difference d


\[\bar{d}=\sum \frac{d_i}{n}= \frac{2+3+1+0+4+2}{6}= 2\]



Standard Deviation:           Sx=(xi-x)2n-1      

\[
\begin{array}{|c|c|c|c|} \hline
Person&di&di-d&(di-d)2\\ \hline
A&2&0&0\\ \hline
B&3&1&1\\ \hline
C&1&-1&1\\ \hline
D&0&-2&4\\ \hline
E&4&2&4\\ \hline
F&2&0&0\\ \hline
\end{array} 
\]

Sx=(di-d)26-1=( 0 + 1 +1 +4 +4 +0)6-1

Sd=105=2 

Now we are ready to perform our hypothesis test.


\section{Step1 : Formally state the null and alternative hypotheses}

\begin{itemize}
\item d : true difference in weight before and after the diet regime 

\item Null Hypothesis               Ho:d = 0        True difference is zero

\item Alternative Hypothesis      Ha:d 0       True difference is not zero 

\item Remark: This is a two tailed test
\end{itemize}



\section{Step 2 : Compute the test statistic.}


Remember the general structure of a test statistic

TS =Observed Value-Null ValueStd. Error 



From the formulae

We have to compute the standard error for a sample mean. 

( From formulae at back of exam paper)

S.E.(d) =Sdn=26= 0.5773 





\section{Step 3 :  Determine the Critical Value}

\begin{itemize}
\item Small sample (group is less than 30).
(Population variance is unknown.)

\item Use t distribution with n-1 degrees of Freedom.

\item We use Murdoch Barnes Table 7 (Student T distribution)

\item Significance levels is 1\%.  This is a two tailed test.

\end{itemize}

[Blackboard]


Critical Value is 2.571


Step 4 : Decision


[Blackboard]

Is the test statistic in the acceptance region or the rejection region?

It is in the rejection region. We reject the null hypothesis. The diet does work.

%==============================================================%
\subsubsection{Worked Example paired T test}






\end{document}
